---
title: "Whaling"
author: "Danielle Moffat"
date: "2024-08-27"
output: html_document
---

# Adding tidyverse and ggthemes
```{r}
library(tidyverse)
library(ggthemes)
```


# Now let's import our whaling dataset
# Note the ".." code means going 'up' in the working directory
```{r}
getwd()
Whaling <- read_csv("../../Reconstruction Assignment/Data/Whaling.csv")
View(Whaling)
```


# Wrangle
```{r}
Whaling <- 
  Whaling |>
  mutate(Scientific_whaling = `"Scientific Whaling"` - `Commercial Whaling`,
         Commercial_whaling = `Commercial Whaling` - `Traditional Whaling`, 
         Traditional_whaling = `Traditional Whaling`, .keep = "unused")

``
```


# Command Option i is a shortcut for a ```{r} chunk
# Now I am uploading the data (this is only required if the dataset from excel was unable to import into R studio)
```{r}
# Load necessary libraries
library(magrittr)
library(ggplot2)
library(dplyr)



# Melt the data for plotting
Whaling_long <- Whaling %>%
  pivot_longer(cols = c(Scientific_whaling, Traditional_whaling, Commercial_whaling), 
               names_to = "Type", 
               values_to = "Mortality")

# Preview the data
Whaling_long
```

```{r}
# Create a line graph
ggplot(Whaling_long, aes(x = Year, y = Mortality, color = Type, group = Type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Whale Mortality by Type of Whaling",
    x = "Year",
    y = "Individuals Mortality",
    color = "Type of Whaling",
    caption = "Source: IWC"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.title.x = element_text(hjust = 0.5),  # Center the x-axis label
    axis.title.y = element_text(hjust = 0.5),   # Center the y-axis label
  ) +
  scale_y_continuous(
    limits = c(0, 1500),  # Set Y-axis from 0 to 2500 by creating a vector
    breaks = seq(0, 1500, 500)  # Define breaks every 500 units for better readability
  )
```

# ALternatively

```{r}
# Create a line graph
Whaling %>%
  pivot_longer(cols = c(Scientific_whaling, Traditional_whaling, Commercial_whaling), 
               names_to = "Type", 
               values_to = "Mortality") |>
ggplot(aes(x = Year, y = Mortality, color = Type, group = Type)) +
  #geom_line(size = 1) +
  geom_point(size = 2) +
  geom_smooth() +
  labs(
    title = "Whale Mortality by Type of Whaling",
    x = "Year",
    y = "Individuals Mortality",
    color = "Type of Whaling",
    caption = "Source: IWC"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.title.x = element_text(hjust = 0.5),  # Center the x-axis label
    axis.title.y = element_text(hjust = 0.5),   # Center the y-axis label
  ) +
  scale_y_continuous(
    limits = c(0, 1500),  # Set Y-axis from 0 to 2500 by creating a vector
    breaks = seq(0, 1500, 500)  # Define breaks every 500 units for better readability
  )
```